\sysfull

### 2017-07-31 paper/background.tex
revised figure to have WAR bug

Software has different behavior when executed intermittently than when executed
with continuous power.  Each power failure clears volatile memory (registers,
stack, and global variables). Non-volatile memory (e.g., FRAM) persists across
power failures. On a power failure, control flows back to a prior point in the
execution. By default, after a power failure control flows to the beginning of
{\tt main()}; a host of recent work proposed variations of the intermittent
execution model that restart execution differently. Some systems restore to a
checkpoint of execution context (i.e., registers and stack) stored in
non-volatile memory~\cite{mementos,hibernusplusplus,quickrecall,idetic}.

@@

Data may become inconsistent if an attempt to execute some computation {\em
writes} to a non-volatile variable, then power fails, then a second attempt to
re-execute the same computation {\em reads} the value written in the first
attempt.  Figure~\ref{fig:code_demo_incosistency} illustrates how state can
become inconsistent in an intermittent execution using cyclic redundancy check
(CRC) code from MIBench~\cite{hicks_mibench2_2016}. The code computes CRC for
an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder
$\texttt{Rmnd}$. Even with programmer defined checkpoints (i.e., line 3 in
Fig.~\ref{fig:crc_example}) of registers, stack, and non-volatile variables
(labeled \texttt{NV}), the code may compute  $\texttt{data}$ incorrectly.  The
code includes a write after read (WAR) hazard of variable $\texttt{Rmnd}$.  WAR
hazard is caused by consecutive write and read in-between two consecutive
checkpoints using $\texttt{checkpoint()}$ function: $\texttt{data}$ variable is
written \emph{before} (line 10 in Fig.~\ref{fig:crc_example_war})
$\texttt{Rmnd}$ is read (line 6 in Fig.~\ref{fig:crc_example_war}), or more
specifically---updated before a new execution loop.

>>>>

Software has different behavior when executed intermittently than when executed
with continuous power.  Each power failure clears volatile memory (registers,
stack, and global variables). Non-volatile memory (e.g., FRAM) persists across
power failures. On a power failure, control flows back to a prior point in the
execution. By default, after a power failure control flows to the beginning of
{\tt main()}. Early work on intermittent computing preserve intermittent
progress by periodically capturing and restoring to a non-volatile checkpoint
of the volatile execution context (i.e., registers and
stack)~\cite{mementos,hibernusplusplus,quickrecall,idetic}

@@

Checkpoints of volatile state preserve intermittent progress, but do not ensure
data consistency~\cite{dino,chain,ratchet}.  Data may become inconsistent if an
attempt to execute some computation {\em writes} to a non-volatile variable,
then power fails, then a second attempt to re-execute the same computation
incorrectly {\em reads} the value written in the first attempt, rather than the
variable's original value.  The situation occurs when code includes a write
after read (WAR) dependence between operations that manipulate non-volatile
variables~\cite{ratchet,dino,alpaca}.

Figure~\ref{fig:code_demo_incosistency} illustrates how state can become
inconsistent in an intermittent execution using cyclic redundancy check (CRC)
code from MIBench~\cite{hicks_mibench2_2016}. The code computes CRC for an
$\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
Even with the checkpoint at line 3 in Fig.~\ref{fig:crc_example}, the code may
compute  $\texttt{data}$ incorrectly because  of a WAR dependence between the
read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.  
Figure~\ref{fig:crc_example_war} shows an execution of the code in which
power fails during the second loop iteration

 $\texttt{Rmnd}$ is read (line 6 in Fig.~\ref{fig:crc_example_war}), or more
specifically---updated before a new execution loop.

### 2017-07-31 paper/texFigCode/crc_example_war.tex
revised figure to have WAR bug

\begin{lstlisting}[style = mystyle]
$\text{\textcolor{red}{checkpoint()}}$; //Iteration 1: correct execution of the loop body
data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
Rmnd = crcTbl[data] ^ (Rmnd << 8);

$\text{\textcolor{red}{checkpoint()}}$; //Iteration 2: break before reminder calculation
data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
$\text{\textcolor{red}{\lightning\hspace{0.5cm}\lightning\hspace{0.5cm}\lightning}\hspace{0.5cm}}\textcolor{red}{\circlearrowright}$ //Power interrupt and restart

$\text{\textcolor{red}{checkpoint()}}$; //Iteration 3: wrong/stale reminder used for data
data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
$\ldots$
\end{lstlisting}

>>>>

\begin{lstlisting}[style = mystyle]
$\text{\textcolor{red}{checkpoint()}}$; //Correct execution iteration 1 
data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
Rmnd = crcTbl[data] ^ (Rmnd << 8);
$\text{\textcolor{red}{checkpoint()}}$; //Capture checkpoint before iteration 2
data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
Rmnd = crcTbl[data] ^ (Rmnd << 8);      
$\text{\textcolor{red}{\lightning\hspace{0.5cm}\lightning\hspace{0.5cm}\lightning}\hspace{0.5cm}}$ //Power fails after Rmnd written, before checkpoint()

$\textcolor{red}{\circlearrowright}\text{\textcolor{red}{checkpoint()}}$; //Restart iter. 2 from checkpoint; $\text{\textcolor{red}{Rmnd is stale.}}$
data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8)); 
$\textcolor{red}{\times}\text{\textcolor{red}{ERROR: data updated with stale value of Rmnd}}$

$\ldots$
\end{lstlisting}
